---
title: "Midterm BANA7038"
author: "Garima Sharma"
date: "November 27, 2015"
MNumber: M08586245
output: html_document
---

####Summary

In this module, we are analyzing the data provided for Aircraft (with number of observations = 800). 

Firstly, we are reading the data, calculating sample size, cleaning the data based on the conditions provided for height, speed_ground, speed_air, duration and distance, visualizing the data, selecting the parameters, model checking, re-modeling with higher degree, determining model adequacy and checking goodness of fit.

Based on the various functions and tests performed on the data, we can conclude that the model is a best fit and landing distance is dependent on the make of aircraft, ground speed and height over threshold.

Analysing the model,

y = $\beta_0$ + $\beta_1$x1 + $\beta_2$x2 + $\beta_3$x2 + $\beta_4$x3^2

where $\beta_i$ are the co-efficients which are represented as:

y(distance) = 1774.48 + 402.75 * x1 + 13.71 * x2 - 68.82 * x3 + 0.69 * $x3^2$   

x1 = Type of aircraft with Boeing = 1 and Airbus = 0

x2 = height

x3 = speed_ground

**1. Reading the data**

#####Clean Memory and Read Data
```{r}
rm(list=ls())
landingData = read.csv("landing.csv", header = TRUE)
```

To read data from csv file into dataset I have used the function read.csv() and imported data to landingData.

**2. Data Exploration**

#####Calculating Sample Size
```{r}
n = dim(landingData)[1]
variables = names(landingData)
```

Sample size can be obtained using the above parameters. We can see that no. of observation in our sample is `r n` and names of the variables are `r variables`.

#####Visualizing Data

From the data, we can observe that there is a linear relationship between variables 'speed_ground' and 'speed_air' with variable 'distance'.
```{r echo=FALSE}
pairs(landingData,pch=20)
```

We can further analyze the numeric variables of landing data by plotting their histograms.
```{r, echo=FALSE, fig.width=12, fig.height=6}
landingData <- read.csv("C:/Garima/2nd Flex/DAM/MidTerm/landing.csv", header=TRUE)
par(mfrow=c(2,4))
hist(landingData$duration, 
     main="Histogram for Flight Duration", 
     xlab="Flight Duration", 
     border="blue", 
     col="green")
hist(landingData$no_pasg, 
     main="Histogram for No. of Passengers", 
     xlab="Passengers", 
     border="blue", 
     col="green")
hist(landingData$speed_ground, 
     main="Histogram for Speed Ground", 
     xlab="Speed Ground", 
     border="blue", 
     col="green")
hist(landingData$speed_air, 
     main="Histogram for Speed Air", 
     xlab="Speed Air", 
     border="blue", 
     col="green")
hist(landingData$height, 
     main="Histogram for Height", 
     xlab="Height", 
     border="blue", 
     col="green")
hist(landingData$pitch, 
     main="Histogram for Pitch", 
     xlab="Pitch", 
     border="blue", 
     col="green")
hist(landingData$distance, 
     main="Histogram for Distance", 
     xlab="Distance", 
     border="blue", 
     col="green")
```

We can observe the data further using the summary function. It can be seen that there are two types of aircrafts i.e. Airbus and Boeing. 

Speed_Air has 600 NA's.

```{r echo=FALSE}
summary(landingData)
```

#####Cleaning the data

According to question, I will consider the values for normal flights.

For duration, normal flight should always be greater than 40min.

```{r}
landingData1 = landingData[ (landingData$duration > 40), ]
```

For normal landing, 140mph >= speed_ground >= 30mph and 140mph >= speed_air >= 30mph.

```{r}
landingData2 = landingData1[(landingData1$speed_ground >= 30), ]
landingData3 = landingData2[(landingData2$speed_ground <= 140), ]
landingData4 = landingData3 [!is.null(landingData3$speed_air >= 30),]
landingData5 = landingData4 [!is.null(landingData4$speed_air <= 140),]
```

Height of the aircraft must be at least 6 meters high.

```{r}
landingData6 = landingData5[(landingData5$height >= 6), ]
```

Length of an airport runway is less than 6000 feet.

```{r}
myLandingData = landingData6[(landingData6$distance < 6000), ]
```

Cleaned data is myLandingData and no. of observations are `r dim(myLandingData)[1]`.

**3. Visualizing Data**

We can again visualize data using pairs() function and observe that there is still exist a linear relationship between variables 'speed_ground' and 'speed_air' with variable 'distance'.

```{r echo=FALSE}
pairs(myLandingData)
```

It can be seen that using the variables aircraft, speed_ground, speed_air and height, we can derive distance.

```{r echo=FALSE}
attach(myLandingData)
model <- lm(distance ~ aircraft + duration + no_pasg + speed_ground + speed_air  + height + pitch)
summary(model)
```

**4. Selecting the variables**

1. For variable Aircraft,

```{r fig.width=3, fig.height=3} 
aircraftType = rep(1,length(aircraft)) * (aircraft == "boeing")
plot(aircraftType, distance)
```

After plotting the aircraft type vs distance, we can observe that mean distance of aircraft boeing is higher than the mean distance of aircraft airbus.

2. For variable Height,

We will keep height variable as we can observe from least square fitted model that height has high importance.

3. For variable Speed_ground,

This variable also has utmost importance as it varies linearily with the variable distance.

4. For variable Speed_Air,

Since this variable has 600 NAs, so we can analyze our data to find if there is some other variable which can act as its substitute.

Degree of co-relation between 'speed_air' and other variables can be observed by:

```{r fig.width=3, fig.height=3} 
cor(myLandingData[,2:7],use="pairwise.complete.obs")
```

We can see from correlation coefficient of 0.99 between speed_air and speed_ground that they both vary linearily.

We can plot these variables for better understanding:

```{r fig.width=3, fig.height=3} 
plot(speed_ground, speed_air)
```

It can be observed that speed_air varies linearily with speed_ground for speed_ground > 90.

**5. Model Checking**

#####New Linear Regression Model for the above variables:

```{r echo=FALSE}
newModel <- lm(distance ~ aircraftType + height + speed_ground)
summary(newModel)
```

Plotting the regression for each independent variable:

```{r echo=FALSE, fig.width=7, fig.height=3}
par(mfrow=c(1,3))
plot(distance~aircraftType)
plot(distance~height)
plot(distance~speed_ground)
```

We can observe from the plots that speed_ground is not completely linear with distance.

To verify it further, we can generate residual plots:

```{r echo=FALSE, fig.width=7, fig.height=3}
par(mfrow=c(1,4))
residual = newModel$res 
plot(residual~aircraftType)
plot(residual~height)
plot(residual~speed_ground)
plot(residual~newModel$fitted.values)
```

It can be observed from residual plots that speed_ground doesn't hold a linear relationship with distance. When the value of speed_ground increases above 80, there exist a linear relationship however, for the complete values between 40mph to 80mph, there exist a quadritic relationship.

**Re-Modeling**

We can further perform the re-modeling by adding a higher degree variable in the model mentioned above. 

```{r echo=FALSE, fig.width=7, fig.height=3}
speedGroundSq = I(speed_ground^2)
modelLatest <- lm(distance ~ aircraftType + height + speed_ground + speedGroundSq)
summary(modelLatest)
```

Residual Plots:

```{r echo=FALSE, fig.width=7, fig.height=3}
par(mfrow=c(1,5))
speedgroundsquare = speed_ground^2
residualLatest = modelLatest$res
plot(residualLatest~aircraftType)
plot(residualLatest~height)
plot(residualLatest~speed_ground)
plot(residualLatest~speedgroundsquare)
plot(residualLatest~modelLatest$fitted.values)
```

**7. Model Adequacy**

#####7.1 Plotting Residuals of other parameters
```{r , fig.width=7, fig.height=3}
par(mfrow=c(1,4))
plot(residualLatest~pitch)
plot(residualLatest~no_pasg)
plot(residualLatest~duration)
```

As we can observe that, non-model parameters do not exhibit any specific pattern hence it demonstrates that the model is a good fit.

#####7.2 Calculating coefficient of Determination

```{r echo=FALSE}
rSq = summary(modelLatest)$r.square
```

A high value of R^2 of `r rSq` demonstrates that the model is a good fit.

#####7.3 Performing T Test
While performing T Test, we can reject the null hypothesis as the t values of all regressors are less than the value of tcritical.

Also since the calculated p-values are less than $\alpha = 0.05$ so it again proves that we can reject the null hypothesis.

Hence we can conclude that 'distance' demonstrates a relationship with all the variables.

#####7.4 Performing F Test

Calcuating SST, SSRES, SSR and F
```{r} 
meanDis = mean(distance)
SST=sum((distance-meanDis)^2)
SSRes=sum((distance-modelLatest$fitted.values)^2)
SSR=sum((modelLatest$fitted.values-meanDis)^2)
fVal=SSR/2/(SSRes/(n-2-1))
```

SST(Total) = `r SST`
SSRes(Residual) = `r SSRes`
SSR(Regression) = `r SSR`

Observing that value of FStatistic = `r fVal` > FCritical, null hypothesis H0 can be rejected which shows that 'distance' has a relationship with all the variables.

#####7.5 Calculating Multicolinearity

Calculating multicolinearity without speed_ground^2:

```{r}
modelN <- lm(distance ~ aircraftType + height + speed_ground)
library(car)
vif(model4)
```

Since, the value of VIF is near 1 it means that there is no multicolinearity between the regressors.

**8. Conclusion**

####We can conclude from the above observations that this model is the best fit model due to the following reasons:

1. All the variables plays a significant role in this model.

2. Resiual plots do not follow a particular pattern and is distributed randomly.

3. P value of 2.2e-16 is very small and less than 0.05.

4. R^2 of 0.977 is very near to 1 hence the model can determine 97.7% of the variability in distance.

5. Landing distance is dependent on the make of aircraft, ground speed and height over threshold.

