---
title: "Assignment 2"
author: "Garima Sharma"
date: "November 9, 2015"
output: html_document
---

Assignment 2

####### Question 1 ###################### 

To read rocket propellent data, I have used read.delim().

```{r}
rm(list=ls())
rocketdata <- read.delim("Data-ex-2-1 (Rocket Prop).txt",h=T)
rocketdata
```


####### Question 2 ###################### 

I have plotted the rocket propellant data as follows:

```{r}
names(rocketdata)
print(rocketdata)
n=dim(rocketdata)[1]
```

```{r, echo=FALSE}
plot(rocketdata$x,rocketdata$y,pch=20, main="Rocket Propellant Data", col.main="red", font.main=4, xlab="Shear Strength", ylab="Age of Propellant", col.lab=rgb(0,0.5,0))
```

####### Question 3 ###################### 

I have performed linear regression without using lm in the following manner:

3.1 

```{r}
meanx = mean(rocketdata$x)
meany = mean(rocketdata$y)
Sxx=sum((rocketdata$x-meanx)^2)
Sxy=sum((rocketdata$x-meanx)*rocketdata$y)
```

Value of Sxx is `r Sxx` and value of Sxy is `r Sxy` .

3.2 

```{r}
beta1 = Sxy / Sxx
beta1
beta0 = meany - beta1*meanx
beta0
```

Value of $\beta0$ is `r beta0` and value of $\beta1$ is `r 

3.3 

```{r}
SST = sum((rocketdata$y - meany)^2)
SST
yrocketdata = beta1 * rocketdata$x + beta0
yrocketdata
SSR = sum((yrocketdata - meany)^2)
SSR
SSRES = sum((rocketdata$y - yrocketdata)^2)
SSRES
```

Value of SST is `r SST`, value of SSR is `r SSR` and value of SSRES is `r SSRES`.

3.4 

```{r}
rocketdataRsquare = SSR / SST
```

Value of R^2 is `r rocketdataRsquare`.

####### Question 4 ###################### 

I have used the following method to calculate linear regression using lm().

4.1 

```{r}
modelrocketdata = lm(y ~ x, data=rocketdata)

sxxnew=sum((rocketdata$x-meanx)^2)
sxynew=sum((rocketdata$x-meanx)*rocketdata$y)
```

Value of Sxx is `r sxxnew` and value of Sxy is `r sxynew` .

4.2 

```{r}
beta0new = modelrocketdata$coefficients[1]
beta1new = modelrocketdata$coefficients[2]
```

Value of $\beta0$ is `r beta0new` and value of $\beta1$ is `r beta1new`.

4.3 

```{r}
sstnew = sum((rocketdata$y - meany)^2)
ssrnew = sum((modelrocketdata$fitted.values - meany)^2)
ssresnew = sum((rocketdata$y - modelrocketdata$fitted.values)^2)
```

Value of SST is `r sstnew`, value of SSR is `r ssrnew` and value of SSRES is `r ssresnew`.

4.4 

```{r}
rocketdataRsquarenew = summary(modelrocketdata)$r.square
```

Value of R^2 is `r rocketdataRsquarenew`.

####### Question 5 ###################### 

Hypothesis Testing and Interval Estimation

5.1 t Test

t-statistic is calculated as $t0 = \beta1 / SE(\beta1)$

t-statistic can be calculated using summary of data read from the text file above
and calculating linear regression.

```{r}
modelrocketdata = lm(y ~ x, data=rocketdata)
summary(modelrocketdata)$coef
tvalue = summary(modelrocketdata)$coef[,1]/summary(modelrocketdata)$coef[,2]
tcritical = qt(1-(0.05/2), 18)
```

t-statistic calculated is `r tvalue`.

Assuming alpha as `r 0.05`, t critical can be calculated as `r tcritical`.

I can observe that value if t-statistic is less than the value of t-critical, hence null hypothesis is rejected. 

I can conclude that there exists a linear relationship between y and x.

Using p-value as calculated below which is also obtainable from summary data,

```{r}
pval = 2*(1-pt(abs(summary(modelrocketdata)$coef[,1]/summary(modelrocketdata)$coef[,2]),18))
```

Since, pvalue of `r pval` is less than the value of 0.05, we can reject the null hypothesis.

5.2 ANOVA Test

```{r}
MSRES=SSRES/(n-2)
fvalue = round((SSR/1)/(SSRES/(n-2)),2)
```

```{r}
n=20
alpha = 0.05
anova(modelrocketdata)
F0 = round((SSR/1)/(SSRES/(n-2)),2)
crticalf = qf(1-alpha,1,n-2)
```

Value of F0 is `r fvalue` which can also be calculated from anova method.
value of fcritical is `r crticalf`.

Comparing the value of F statistic which is greater than the F critical, we can conclude that null hypothesis can be rejected. Hence there is a linear relationship between y and x.

F test p-value is `1-pf(fvalue, 1, 18)`.

p-value is `r 1-pf(fvalue, 1, 18)`

Since p-value is less than the value of alpha `r alpha`, we can reject the null hypothesis. 

5.3 Calculating Confidence Interval

5.3.1 I am calculating confidence interval for coefficients by using the following formulas for slope as well as intercept:

```{r}
alpha = 0.05
n=20
modelrocketdata = lm(y ~ x, data=rocketdata)
beta0hat = modelrocketdata$coefficients[1]
beta1hat = modelrocketdata$coefficients[2]
tcritical = qt(1-alpha/2, n-2)
sebeta0hat = summary(modelrocketdata)$coef[1,2]
sebeta1hat = summary(modelrocketdata)$coef[2,2]
```

Confidence Interval at `r 100*(1-alpha)`% level of significance

For $\beta0$, 
`r round((beta0hat - tcritical*sebeta0hat),3)` <= $\beta0$ <= `r round((beta0hat + tcritical*sebeta0hat),3)`

For $\beta1$,
`r round((beta1hat - tcritical*sebeta1hat),3)` <= $\beta1$ <= `r round((beta1hat + tcritical*sebeta1hat),3)`

5.3.2 Calculating for fitted values, 

I calculated the fitted values and confidence interval of y by using the given value of x as mentioned below:

```{r echo = FALSE}

confidence1= predict(modelrocketdata,interval = c("confidence"),level =0.95, type="response")
msres=SSRES/(n-2)
merr = tcritical * sqrt ((MSRES)*(((1/n) + ((rocketdata [,2]-mean(rocketdata [,2]))^2)/Sxx)))

fittedy = beta0hat + beta1hat * rocketdata[,2]
ylower = fittedy - merr
yupper = fittedy + merr
W = yupper - ylower

#confintervaldata = data.frame(fittedy, ylower, yupper,W)
#cdata = cbind(rocketdata, confintervaldata)

#cdata

```

5.3.3 We can calculate predicted value as:

```{r echo = FALSE}

xpredictedval = seq(0,40)
confidenceval= predict(modelrocketdata,newdata=data.frame(x=xpredictedval),interval = c("confidence"),level =0.95, type="response")
confidenceval

#marginerr = tcritical * sqrt ((MSRES)*  ((1)+ (1/n) + (((xpredictedval-mean(xpredictedval))^2)/sxx)))
#ypredictedval = beta0hat + beta1hat * xpredictedval
#ylower = ypredictedval - marginerr
#yupper = ypredictedval + marginerr
#Width = yupper - ylower
#cipred = data.frame(xpredictedval, ypredictedval, ylower, yupper,Width)
#cipred

```

5.4 Plot data points, the regression line, the confidence interval for fitted values and the confidence interval for predict values 

```{r echo=FALSE}
plot(rocketdata$x,rocketdata$y,pch=20)
abline(modelrocketdata, col="red")
lines(rocketdata$x, ylower,col="green",lty=2)
lines(rocketdata$x, yupper,col="green",lty=2)


plot(rocketdata$x,rocketdata$y,pch=20)
modelfinal = lm(y~x, data = rocketdata)
abline(modelrocketdata, col="red")
lines(xpredictedval,confidenceval[,2],col="blue",lty=2)
lines(xpredictedval,confidenceval[,3],col="blue",lty=2)

```
